\bf Approximating Likelihood Ratios with Calibrated Discriminative Classifiers

### 2016-01-22 jasa/paper-jasa.tex
Section 6: related work

@@

\section{Related works}
\label{sec:related}

\glnote{Add related works on density ratio estimation.}
\glnote{Remove others that are less directly related.}

\cite{ClaytonScott} and \cite{JMLR:v14:tong13a} consider the machine learning
problem associated to Neyman-Pearson hypothesis testing. As in this work, they
consider the situation where one does not have access to the underlying
distributions, but only has i.i.d. samples from each hypothesis. This work
generalizes that goal from the Neyman-Pearson setting to generalized likelihood
ratio tests and emphasizes the connection with classification. Perhaps a  formal
treatment similar to the Neyman-Pearson case can be brought to bear in this more
general setting. In a similarly titled work, \cite{Gutmann2014} advocate using
the cross-validated classification accuracy as the similarity metric used in
ABC. While the goal there is also parameter inference in the likelihood-free
setting,  ``classifier ABC'' is very different than the approach presented here.
\cite{TommiJaakkola} explore a way of leveraging generative models to derive
kernel functions for use in discriminative methods. This interesting work is
distinct from the point made here in which the generative model is being used
for the purpose of providing training data and calibration.
%\cite{McCallum} consider a hybrid generative/discriminative classifier; however, the goal of that work is not to leverage a generative model for the data, but to use both approaches to learn different subsets of the parameters in a single hybrid classifier.
\cite{BiancaZadrozny} emphasize the importance of calibrated probability
estimates from decision trees and naive Bayesian classifiers and investigate
various approaches to achieve this. In contrast to that work, we are not
interested in calibrated probability estimates for $p(y|x)$ for individual
events, but instead we use the calibration to correct for non-linear
transformations of the target likelihood ratio and, perhaps, to provide
calibrated p-values based on those likelihood ratio tests. \cite{Ihler2004} take
on a different problem (tests of statistical independence) by using machine
learning algorithms to find  scalar maps from the high-dimensional feature space
that achieve the desired statistical goal when the fundamental high-dimensional
test is intractable.

\cite{Neal:2007zz} also considered the problem of approximating the likelihood
function when only a generative model is available. That work sketches a scheme
in which one uses a classifier with both $x$ and $\theta$ as an input to serve
as a dimensionality reduction map. The key distinction comes in the handling of
$\theta$.  Neal says  ``we cannot use the classifier on real data, since we
don't know the correct value for [$\theta]$'' and goes on to outline an approach
where one uses regression on a per-event basis to estimate $\hat{\theta}(x)$ and
perform the composition $s(x; \hat{\theta}(x))$ (much like
profiling).\footnote{Neal considers a lower-dimensional `bottlenecks'
$\theta^*=g(\theta)$, which are not essential to the discussion here.}  This can
lead to a significant loss of information since (at least in most particle
physics examples) a single event carries little information about the true value
of $\theta$, though the full data set $D$ may be informative -- for instance, a
single observation would not be sufficient to estimate the variance of a
distribution, though repeated observations would.  The work of Neal correctly
identifies this as an approximation of the target likelihood even in the case of
a ideal classifier. In contrast, the approach described here does not eliminate
the dependence of the classifier on $\theta$.\footnote{As a technical point,
in Neal's work, the focus is on approximating the likelihood function (up to a
multiplicative constant), which is equivalent to evaluating the ratio  with
respect to a fixed $\theta_1$ as in Eq.~\ref{eq:mle_withs}. In Neal's case, the
dependence on $\theta$ is eliminated via $s(x; \hat{\theta}(x))$ and the map is
constant; however, in this approach the map ratio is explicitly parameterized in
terms of $\theta$, so the ratio is important for canceling the corresponding
Jacobian factors.} Instead, we embed a parameterized classifier into the
likelihood and postpone the evaluation of the classifier to the point of
evaluation of the likelihood when $\theta$ is explicitly being tested. This
avoids the loss of information that occurs from the regression step
$\hat{\theta}(x)$ proposed by Neal and leads to Theorem~1, which is an exact
result in the case of an ideal classifier. In both cases, the quality of the
classifier is factorized from the calibration of its density, which allows for
valid inference even if there is a loss of power due to a non ideal classifier.

>>>>

\glnote{Add reference to \cite{BiancaZadrozny} here.}

@@

\section{Related work}
\label{sec:related}

The closest work to the proposed method is due to \cite{Neal:2007zz}, who
similarly considers the problem of approximating the likelihood function when
only a generative model is available. That work sketches a scheme in which one
uses a classifier with both $\mathbf{x}$ and $\theta$ as an input to serve as a
dimensionality reduction map. The key distinction comes in the handling of
$\theta$.  Neal argues that a classifier cannot be used on real data, since we
do not know the correct value for $\theta$, and goes on to outline an approach
where one uses regression on a per-event basis to estimate
$\hat{\theta}(\mathbf{x})$ and perform the composition $s(\mathbf{x};
\hat{\theta}(\mathbf{x}))$. As pointed out by the author, this can lead to a
significant loss of information since a single observation $\mathbf{x}$ may
carry little information about the true value of $\theta$, though a full data
set ${\cal D}$ may be informative -- for instance, a single observation would
not be sufficient to estimate the variance of a distribution, though repeated
observations would. The work of \cite{Neal:2007zz} correctly identifies this as
an approximation of the target likelihood even in the case of a ideal
classifier. In contrast, the approach described here does not eliminate the
dependence of the classifier on $\theta$\footnote{As a technical point, in
\citep{Neal:2007zz}, the focus is on approximating the likelihood function (up
to a multiplicative constant), which is equivalent to evaluating the ratio  with
respect to a fixed $\theta_1$ as in Eqn.~\ref{eq:mle_withs}. In
\citep{Neal:2007zz}, the dependence on $\theta$ is eliminated via $s(\mathbf{x};
\hat{\theta}(\mathbf{x}))$ and the map is constant; however, in this approach
the map ratio is explicitly parameterized in terms of $\theta$, so the ratio is
important for canceling the corresponding Jacobian factors. \glnote{Shall we
keep this footenote? Would be better to give details in the text. JASA
recommends not using footnotes.}}. Instead, we embed a parameterized classifier
into the likelihood and postpone the evaluation of the classifier to the point
of evaluation of the likelihood when $\theta$ is explicitly being tested. This
avoids the loss of information that occurs from the regression step
$\hat{\theta}(\mathbf{x})$ proposed by \cite{Neal:2007zz} and leads to
Theorem~\ref{thm:ratio-equivalence}, which is an exact result in the case of an
ideal classifier. In both cases, the quality of the classifier is factorized
from the calibration of its density, which allows for valid inference even if
there is a loss of power due to a non ideal classifier. \glnote{We could shorten this paragraph.}

Also close to our work, \cite{ClaytonScott} and \cite{JMLR:v14:tong13a} consider
the machine learning problem associated to Neyman-Pearson hypothesis testing. In
a similar setup, they consider the situation where one does not have access to
the underlying distributions, but only has i.i.d. samples from each hypothesis.
This work generalizes that goal from the Neyman-Pearson setting to generalized
likelihood ratio tests and emphasizes the connection with classification.
% Perhaps a  formal treatment similar to the Neyman-Pearson case can be brought to
% bear in this more general setting.
\cite{Ihler2004} take on a different problem
(tests of statistical independence) by using machine learning algorithms to find
scalar maps from the high-dimensional feature space that achieve the desired
statistical goal when the fundamental high-dimensional test is intractable.
\glnote{Add other references on hypothesis testing with ML?}

% In a similarly titled work, \cite{Gutmann2014} advocate using
% the cross-validated classification accuracy as the similarity metric used in
% ABC. While the goal there is also parameter inference in the likelihood-free
% setting,  their method is very different than the approach presented here.
% \cite{TommiJaakkola} explore a way of leveraging generative models to derive
% kernel functions for use in discriminative methods. This interesting work is
% distinct from the point made here in which the generative model is being used
% for the purpose of providing training data and calibration.
% \cite{McCallum} consider a hybrid generative/discriminative classifier;
% however, the goal of that work is not to leverage a generative model for the
% data, but to use both approaches to learn different subsets of the parameters in
% a single hybrid classifier.
% \cite{BiancaZadrozny} emphasize the importance of calibrated probability
% estimates from decision trees and naive Bayesian classifiers and investigate
% various approaches to achieve this. In contrast to that work, we are not
% interested in calibrated probability estimates for $p(y|x)$ for individual
% events, but instead we use the calibration to correct for non-linear
% transformations of the target likelihood ratio and, perhaps, to provide
% calibrated p-values based on those likelihood ratio tests.

More generally, likelihood ratio testing directly relates to the density ratio
estimation problem, which consists in estimating the ratio of two probability
densities from given and finite collections of observations ${\cal D}_0$ and
${\cal D}_1$. Density ratio estimation is connected to many machine learning
fundamental problems, including domain adaptation~\citep{sugiyama2012machine}, probabilistic
classification and regression~\citep{vapnik1998statistical}, outlier
detection~\citep{hido2011statistical}, and many others. In contrast with the
proposed method which decouples reduction from calibration, most of the
approaches proposed within the literature (see
\cite{sugiyama2012density,gretton2009covariate,nguyen2010estimating,vapnik2013constructive}
and references therein) provide solutions for estimating
$r(\mathbf{x};\theta_0,\theta_1)$ directly from $\mathbf{x}$. Under some
assumptions, the convergence of the obtained estimates was also proven for some
of these approaches.
